Model Checking Differentially Private Properties

### 2018-01-27 examples.tex
more changes ;)
threshold example

\begin{example}\label{exa:survey}
  Figure~\ref{figure:2-dp} shows the survey
mechanism and its corresponding Markov chain. In the figure, the states
$+$ and $-$ denote the diagnoses are positive or negative
respectively; the states $Y$ and $N$ denote answers to the survey
question. We have $AP=\{Y,N\}$, $L(Y)=\{Y\}$, $L(N)=\{N\}$, and $L(+)=L(-)=\emptyset$. \lz{draw it on the picture?}

@@

Similar to the survey mechanism, it is
straightfoward to model the truncated $1/2$-geometric
mechanisms as Markov chains. Let the state $in_k$ and $out_l$
denote the input value $k$ and output value $l$ respectively. Define
the state set $S = \{ in_k, out_k : k \in \underline{n} \}$.
The probability transition $\wp (in_k, out_l)$ is the probability of the
output value $l$ on the input value $k$ as defined in the
mechanism. Figure~\ref{figure:geometric-mechanism-markov-chain} gives
the Markov chain of the truncated
$1/2$-geometric mechanism over $\underline{5}$.

\subsection{Markov Chains for Differentially Private Mechanisms}
Now we describe our main ideas for the modelling differentially private mechanisms. 
A state $s$ stores at least the a database $x\in \mathcal{X}^n$, which is the data hold by a trusted \emph{curator}. A transition corresponds to the data analysis mechanisms.
States are typically labelled with the corresponding outputs
the curator generates.

@@

Below we describe an online model.

\begin{figure}
    \centering
    \begin{subfigure}{.4\columnwidth}
\resizebox{\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
    distance=1cm,node/.style={circle,draw}]
    \node[node] (p) at ( -1.5,  0) { $+$ };
    \node[node] (Y1) at ( 0, .8) { $Y_1$ };
    \node[node] (N1) at ( 0, -.8) { $N_1$ };
    \node[node] (Y2) at (1.5,  .8) { $Y_2$ };
    \node[node] (N2) at ( 1.5, -.8) { $N_2$ };

    \path
    (p) edge node [above] { $\frac{3}{4}$ } (Y1)
    (p) edge node [below] { $\frac{1}{4}$ } (N1)
    (Y1) edge node [above] { $\frac{3}{4}$ } (Y2)
    (Y1) edge node [above] { $\frac{1}{4}$ } (N2)
    (N1) edge node [below] { $\frac{3}{4}$ } (Y2)
    (N1) edge node [below] { $\frac{1}{4}$ } (N2)
    ;
    \end{tikzpicture}
}
%        \caption{Markov Chain of }
%        \label{figure:2-dp-mc1}
    \end{subfigure}
    \hspace{.05\columnwidth}
    \begin{subfigure}{.4\columnwidth}
        \resizebox{\columnwidth}{!}{
            \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
    distance=1cm,node/.style={circle,draw}]
    \node[node] (n) at ( -1.5,  0) { $-$ };
    \node[node] (Y1) at ( 0, .8) { $Y_1$ };
    \node[node] (N1) at ( 0, -.8) { $N_1$ };
    \node[node] (Y2) at (1.5,  .8) { $Y_2$ };
    \node[node] (N2) at ( 1.5, -.8) { $N_2$ };

    \path
    (n) edge node [above] { $\frac{1}{4}$ } (Y1)
    (n) edge node [below] { $\frac{3}{4}$ } (N1)
    (Y1) edge node [above] { $\frac{1}{4}$ } (Y2)
    (Y1) edge node [above] { $\frac{3}{4}$ } (N2)
    (N1) edge node [below] { $\frac{1}{4}$ } (Y2)
    (N1) edge node [below] { $\frac{3}{4}$ } (N2)
    ;
            \end{tikzpicture}
        }
%        \caption{Markov Chain of}
%        \label{figure:2-dp-mc2}
    \end{subfigure}
    \caption{Markov Chain of }
    \label{figure:2-dp-mc}
\end{figure}
\begin{example}
  First, assume that the curator allows at most two survey queries. The corresponding model is depicted in Figure~\ref{figure:2-dp-mc}, where states $Y_1,Y_2$ are labelled by $Y$, and $N_1,N_2$ are labelled by $N$.
\end{example}

\lz{tbd, the complex threshold example we discussed, by Depeng}

>>>>

\begin{example}\label{exa:survey}
  Figure~\ref{figure:2-dp} shows the survey
mechanism and its corresponding Markov chain. In the figure, the states
$+$ and $-$ denote the diagnoses are positive or negative
respectively; the states $Y$ and $N$ denote answers to the survey
question. We have $AP=\{Y,N\}$, $L(Y)=\{Y\}$, $L(N)=\{N\}$, and $L(+)=L(-)=\emptyset$. States $+$ and $-$ are neighbors.\lz{draw it on the picture?}

@@

Similar to the survey mechanism, it is
straightfoward to model the truncated $1/2$-geometric
mechanisms as Markov chains.

\begin{example}
Let the state $in_k$ and $out_l$
denote the input value $k$ and output value $l$ respectively. Define
the state set $S = \{ in_k, out_k : k \in \underline{n} \}$.
The probability transition $\wp (in_k, out_l)$ is the probability of the
output value $l$ on the input value $k$ as defined in the
mechanism. Figure~\ref{figure:geometric-mechanism-markov-chain} gives
the Markov chain of the truncated
$1/2$-geometric mechanism over $\underline{5}$. Further, assume
$L(out_i) = \{ out_i \}$ and $\Delta (f) = 1$. Then $in_k$ and $in_l$ are
neighbors iff $| k - l | \leq 1$.
\end{example}

\subsection{Markov Chains for Differentially Private Mechanisms}
Now we describe our main ideas for  modelling differentially private mechanisms.
A state $s$ stores at least the a database $x\in \mathcal{X}^n$, which is the data hold by a trusted \emph{curator}. A transition corresponds to the data analysis mechanisms.
States are typically labelled with the corresponding outputs
the curator generates.

@@

\paragraph{Online Model.}
Below we describe an online model adapted from \cite{}: it .....

We consider the database set $D = \{0,1,2,3,4,5\}$, a threshold
$t \in \{0,1,2,3,4,5\}$, queries $\{f| f $ with sensitivity $\Delta (f) = 1 \}$. In order to protect privacy, a truncated $1/2-$geometric mechanisms over $\underline{5}$ is applied to each query and the threshold, and what we observe is the first time that a query result is greater than or equal to the threshold after interference. We are interested in the differential privacy property of the data in $D$, assuming that they are all neighbours. The modeling can be done in guarded command language and some fragments are shown below.\todo{the algorithm seems redundant}

\begin{algorithm}
    \begin{algorithmic}[1]

        \Function{OnlineModel}{$d$, $t$}

        \State{$\mathbf{If}$  $t=0$ $\rightarrow   2/3 : (\hat{t} = 0) + 1/6 : (\hat{t} = 1) + 1/12 : (\hat{t} = 2)$ }
        \State{$+ 1/24 : (\hat{t} = 3) + 1/48 : (\hat{t} = 4) + 1/48: (\hat{t} = 5);$}
        \State{$| $ $ t=1\rightarrow   1/3 : (\hat{t} = 0) + 1/3 : (\hat{t} = 1) + 1/6 : (\hat{t} = 2)$}
        \State{$+ 1/12 : (\hat{t} = 3) + 1/24 : (\hat{t} = 4) + 1/24: (\hat{t} = 5);$}
        \State{$...$  $...$}
        \Comment{Assignments of $\hat{t}$ by the geometric mechanism}
        \State{$\mathbf{Fi}$}

        \For{each query $f_i$}
        \State{$\mathbf{If}$  $d=0$ $\rightarrow   2/3 : (\hat{f} = 0) + 1/6 : (\hat{f} = 1) + 1/12 : (\hat{f} = 2)$ }
        \State{$+ 1/24 : (\hat{f} = 3) + 1/48 : (\hat{f} = 4) + 1/48: (\hat{f} = 5);$}
        \State{$| $ $ d=1\rightarrow   1/3 : (\hat{f} = 0) + 1/3 : (\hat{f} = 1) + 1/6 : (\hat{f} = 2)$}
        \State{$+ 1/12 : (\hat{f} = 3) + 1/24 : (\hat{f} = 4) + 1/24: (\hat{f} = 5);$}
        \State{$...$  $...$}
        \Comment{Assignments of $\hat{f}$ by the geometric mechanism}
        \State{$\mathbf{Fi}$}
        \State{$\mathbf{If}$ $\hat{f}\ge \hat{t}$ $a_i=\top$ \Return }
        \State{$|$   $\hat{f} < \hat{t}$ $a_i=\bot$}
        \State{$\mathbf{Fi}$}
        \EndFor
        \EndFunction

    \end{algorithmic}
    \caption{Input is a private database $d \in \{0,1,2,3,4,5\}$, an adaptively chosen stream of sensitivity 1 queries $f_1$, . . ., a threshold $t \in \{0,1,2,3,4,5\}$.Output is a stream of answers $a_1$,...}
    \label{algorithm:online-model}
\end{algorithm}

### 2018-01-27 semantics.tex
threshold example

\section{Model Checking Algorithms}
We first describe the model checking algorithm for $\dpCTL$, which follows the classical
algorithms for $\PCTL$ by computing the states satisfying
sub state-formulae inductively~\cite{BK:08:PMC}. It hence suffices to
consider the inductive step where the states satisfying the subformula
$\dpriv{\epsilon}{\delta} (\phi)$ is computed.

>>>>

\subsection{Model Checking Algorithms}
We first describe the model checking algorithm for $\dpCTL$, which follows the classical
algorithms for $\PCTL$ by computing the states satisfying
sub state-formulae inductively~\cite{BK:08:PMC}. It hence suffices to
consider the inductive step where the states satisfying the subformula
$\dpriv{\epsilon}{\delta} (\phi)$ is computed.

### 2018-01-27 specifications.tex
threshold example

\paragraph{Compositional Properties.}
Compositional aspect is one of the building block for differentially private algorithms. According to the compositional theorem \cite{}, for $(\epsilon_1,0)$-differentially private algorithm $M_1$ and
$(\epsilon_2,0)$-differentially private algorithm $M_2$, their combination $M$ defined by $M(x)=(M_1(x), M_2(x))$ is $(\epsilon_1\epsilon_2,0)$-differentially private. The degradation is rooted in the repeated computation. To illustrate this property, we consider the extended example below \lz{here example of the survey example, expanded twice: we should discuss the graph before it is drawn...}

We consider the formula $\dpriv{9}{0} (\X (Y\wedge \X Y))$. A path satisfies $\X (Y\wedge \X Y)$ if the second state satisfies $Y$ and the third state satisfies $Y$ as well. We verify that this formula is satisfied  for state $+$ and $-$. Moreover, the bound $\epsilon=9$ is bound, since \lz{tbd once graph is given}. Finally, the formula $\wedge_{a_1,a_2}\dpriv{9}{0} (\X (a_1\wedge \X a_2))$ corresponds to the desired compositional property for the model, where $a_1,a_2$ range over all atomic propositions $\{Y,N\}$.

@@

Now we consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X (Y\wedge \X Y))$ under the model in ???. Applying the compositional theorem\lz{Bow-Yaw, as we only refer to the book, I wonder whether using the notion $e^\epsilon$ is easier to explain?}, one can choose $\epsilon_2=\epsilon_1^2=1$ and $\delta_2=2\delta_1=1$. However, we can check easily that one can get better privacy parameter $(1,0.5)$ or $(0,\frac{9}{16})$.

>>>>

\paragraph{Compositional Properties.}
Compositional aspect is one of the building block for differentially private algorithms. According to the compositional theorem \cite{}, for $(\epsilon_1,0)$-differentially private algorithm $M_1$ and
$(\epsilon_2,0)$-differentially private algorithm $M_2$, their combination $M$ defined by $M(x)=(M_1(x), M_2(x))$ is $(\epsilon_1\epsilon_2,0)$-differentially private. The degradation is rooted in the repeated computation. To illustrate this property, we consider the extended survey example below where the curator allows two consequent queries.
The corresponding model is depicted in Figure~\ref{figure:2-dp-mc}, where states $Y_1,Y_2$ are labelled by $Y$, and $N_1,N_2$ are labelled by $N$.

\begin{figure}
    \centering
    \begin{subfigure}{.4\columnwidth}
\resizebox{\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
    distance=1cm,node/.style={circle,draw}]
    \node[node] (p) at ( -1.5,  0) { $+$ };
    \node[node] (Y1) at ( 0, .8) { $Y_1$ };
    \node[node] (N1) at ( 0, -.8) { $N_1$ };
    \node[node] (Y2) at (1.5,  .8) { $Y_2$ };
    \node[node] (N2) at ( 1.5, -.8) { $N_2$ };

    \path
    (p) edge node [above] { $\frac{3}{4}$ } (Y1)
    (p) edge node [below] { $\frac{1}{4}$ } (N1)
    (Y1) edge node [above] { $\frac{3}{4}$ } (Y2)
    (Y1) edge node [above] { $\frac{1}{4}$ } (N2)
    (N1) edge node [below] { $\frac{3}{4}$ } (Y2)
    (N1) edge node [below] { $\frac{1}{4}$ } (N2)
    ;
    \end{tikzpicture}
}
%        \caption{Markov Chain of }
%        \label{figure:2-dp-mc1}
    \end{subfigure}
    \hspace{.05\columnwidth}
    \begin{subfigure}{.4\columnwidth}
        \resizebox{\columnwidth}{!}{
            \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
    distance=1cm,node/.style={circle,draw}]
    \node[node] (n) at ( -1.5,  0) { $-$ };
    \node[node] (Y1) at ( 0, .8) { $Y_1$ };
    \node[node] (N1) at ( 0, -.8) { $N_1$ };
    \node[node] (Y2) at (1.5,  .8) { $Y_2$ };
    \node[node] (N2) at ( 1.5, -.8) { $N_2$ };

    \path
    (n) edge node [above] { $\frac{1}{4}$ } (Y1)
    (n) edge node [below] { $\frac{3}{4}$ } (N1)
    (Y1) edge node [above] { $\frac{1}{4}$ } (Y2)
    (Y1) edge node [above] { $\frac{3}{4}$ } (N2)
    (N1) edge node [below] { $\frac{1}{4}$ } (Y2)
    (N1) edge node [below] { $\frac{3}{4}$ } (N2)
    ;
            \end{tikzpicture}
        }
%        \caption{Markov Chain of}
%        \label{figure:2-dp-mc2}
    \end{subfigure}
    \caption{Markov Chain of Double Surveys}
    \label{figure:2-dp-mc}
\end{figure}

We consider the formula $\dpriv{9}{0} (\X (Y\wedge \X Y))$. A path satisfies $\X (Y\wedge \X Y)$ if the second state satisfies $Y$ and the third state satisfies $Y$ as well. We verify that this formula is satisfied  for state $+$ and $-$. Moreover, the bound $\epsilon=9$ is bound, since the probability of satisfying $\X (Y\wedge \X Y)$ from states $+$ and $-$ are $9$ and $1$, respectively. Finally, the formula $\wedge_{a_1,a_2}\dpriv{9}{0} (\X (a_1\wedge \X a_2))$ corresponds to the desired compositional property for the model, where $a_1,a_2$ range over all atomic propositions $\{Y,N\}$.

@@

Now we consider the formula $\dpriv{\epsilon_2}{\delta_2} (\X (Y\wedge \X Y))$ under the model in Figure~\ref{figure:2-dp-mc}. Applying the compositional theorem, one can choose $\epsilon_2=\epsilon_1^2=1$ and $\delta_2=2\delta_1=1$. However, we can check easily that one gets better privacy parameter $(1,0.5)$ or $(0,\frac{9}{16})$.
